### 2025-02-18 paper.tex
Rough pass on Discussion

\documentclass{article}
\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsfonts, bbm}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{placeins}
% \usepackage[hidelinks]{hyperref}

\newcommand{\tsinfer}[0]{\texttt{tsinfer}}
\newcommand{\kwarg}[0]{\texttt{KwARG}}
\newcommand{\argneedle}[0]{\texttt{ARG-Needle}}
\newcommand{\argweaver}[0]{\texttt{ARGweaver}}
\newcommand{\argweaverD}[0]{\texttt{ARGweaver-D}}
\newcommand{\relate}[0]{\texttt{Relate}}
\newcommand{\espalier}[0]{\texttt{Espalier}}
\newcommand{\arbores}[0]{\texttt{Arbores}}
\newcommand{\tskit}[0]{\texttt{tskit}}
\renewcommand{\deg}{\mathrm{deg}}
\newcommand{\comment}[1]{{\it \color{orange} (#1)}}

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% % Captions will be left justified
% \usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
% \renewcommand{\figurename}{Fig}

% adjust inter-paragraph spacing - this one's for you, Gertjan
% \setlength{\parskip}{0.5\baselineskip}

\begin{document}

\linenumbers
\title{Computing likelihoods under the SMC for a general class of ARGs.}

% First authors
\author[1, $\dagger$]{Gertjan Bisschop}
\author[1]{Jerome Kelleher}
% last author
\author[3]{Peter Ralph}

\affil[$\dagger$]{Denotes corresponding author}

\maketitle

\affil[1]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, OX3 7LF, UK}
\affil[2]{Department of Statistics, University of Oxford, OX1 3LB, UK}
\affil[3]{University of Oregon, USA}

\begin{abstract}
Ancestral recombination graphs (ARGs) are the focus of much ongoing research
interest. Recent progress in inference has made ARG-based approaches feasible
across of range of applications, and many new methods using inferred ARGs as
input have appeared. This progress on the long-standing problem of ARG
inference has proceeded in two distinct directions. The first
class of method in which dramatic progress has been made is in the
statistically rigorous inference of ARGs under the Sequentially Markov
Coalescent (SMC), which are now practical for tens-to-hundreds of samples. 
The second class of recently-developed inference methods are based on
approximate models and heuristics which scale to sample sizes two to three
orders of magnitude larger. Although these heuristic methods are reasonably
accurate under many metrics, one significant drawback is that the ARGs they 
estimate do not have the topological properties required to compute a 
likelihood under models such as the SMC under present-day formulations.
In particular, hueristic inference methods typically do not estimate
precise details about recombination events, which are currently
required to compute a likelihood.
In this paper we present a 
novel backwards-time formulation of the SMC 
and derive a straightforward definition 
of the likelihood of a general class of ARG under this model. 
We show that this formulation does not require precise details of recombination events
to be estimated, and is robust to the presence of polytomies. 
We discuss the possibilities for inference that this opens.
\end{abstract}

% This opens up many future directions, most excitingly the possibility
% of combining the strengths of both inference approaches. Applying
% these ideas may allow us to 
% make rigorous inference more scalable, and scalable inference 
% more rigorous.

\section*{Introduction}
Following recent breakthroughs in simulation and inference methods,
there is now intense interest in applying Ancestral Recombination Graphs (ARGs)
to a variety of questions in evolutionary
biology~\citep{lewanski_era_2024,brandt_promise_2024,nielsen_inference_2025}.
ARGs describe the intricately linked paths
of genetic inheritance resulting from
recombination~\citep{hudson_properties_1983,griffiths_ancestral_1996,wong_general_2023},
and contain rich detail about ancestral processes.
% TODO rephrase this a bit, would be a lot of duplication of references
% in the current formulation.
Numerous applications taking advantage of this detailed history
in inferred ARGs are now appearing~\citep{
stern_approximate_2019,
osmond_estimating_2021,
fan_genealogical_2022,
hejase_deep_2022,
guo_recombination-aware_2022,
ignatieva_ongoing_2022,
wang_complex_2022,
zhang_biobank-scale_2023,
nowbandegani_extremely_2023,
ignatieva_distribution_2023,
fan_likelihood_2023,
huang_estimating_2024,
grundler_geographic_2024,
korfmann_simultaneous_2024,
deraje_inferring_2024,
speidel2025high}
and it seems likely that many more will
follow~\citep{harris_database_2019,harris_using_2023}.
Although there are many different approaches to ARG
inference~\citep{wong_general_2023}, two broad classes of method
have been the focus of most recent interest.

% TODO Maybe swap the order of these two paragraphs around, so it's 
% more natrual to talk about the advantages of model-driven stuff next.
The first class of inference methods are statistically rigorous approaches
that sample ARGs under a population genetics model such as the
coalescent with recombination~\citep{hudson_properties_1983},
and its approximatation, the Sequentially Markovian
Coalescent~\citep{mcvean_approximating_2005,marjoram_fast_2006}.
ARGweaver~\citep{rasmussen_genome-wide_2014} is the most popular,
consistently out-performed other methods in terms 
of accuracy in a variety of benchmarks~\citep{brandt2022evaluation}[MORE
citations,anywhere ARGweaver has been tested for accuracy]
and has succesfully been applied in several
contexts~\citep[e.g.][]{de_chimpanzee_2016,shriner_whole_2018,hejase_genomic_2020,
stankowski_genetic_2024}. 
ARGweaver works [by explanation.]
The recently-introduced SINGER method takes a similar approach [doing something 
slightly different]~\citep{deng_robust_2024}.
While SINGER has not been evaluated as thoroughly as ARGweaver yet, it promises
to be even more accurate while being substantially more
scalable~\citep{deng_robust_2024}.

The second class of inference method that has been of recent interest 
is based on heuristics and approximate models.
Relate~\citep{speidel_method_2019},
tsinfer~\citep{kelleher_inferring_2019},
ARG-Needle~\citep{zhang_biobank-scale_2023} 
and Threads~\citep{gunnarsson_scalable_2024}
work on quite different 
principles, but share some common properties. 
Firstly, they are all 
hueristic and approximation driven, favouring computational
scalability over statistical rigor.
Secondly, they all produce a single
ARG as output, inferring a deterministic point estimate with no means 
of quantifying uncertainty. Thirdly (and most importantly for the 
purposes of this paper) the ARGs they produce lack substantial
detail about ancestral recombination events: the estimates are
rough sketches, hopefully capturing large-scale properties, but lacking
in the finer detail of outputs of ARGweaver or SINGER.

There are substantial advantages to the statistically rigorous sampling 
approaches of ARGweaver and SINGER. Firstly, they tend to be more 
accurate; ARGweaver has consistently out-performed other methods in terms 
of accuracy in a variety of benchmarks~\citep{brandt2022evaluation}[MORE
citations,anywhere ARGweaver has been tested for accuracy]
and, while SINGER has not been evaluated as thoroughly yet, it promises
to be even more accurate~\citep{deng_robust_2024}. 
Secondly, the direct connection with
population genetic models greatly aids interpretability, and allows for
natural extensions like the joint estimation of ARGs and
demography~\citep{hubisz_mapping_2020}. Thirdly, sampling ARGs from a
posterior distribution provides a means of quantifying the uncertainty 
around estimates, which can be used in downstream applications [citation
of this?].

Despite these significant advantages of rigorous inference, however,
hueristic methods remain of interest because of the profound difference
in computational scalability. Although SINGER is a major step forward
in terms of scalability over ARGweaver and other methods, inference is still 
only feasible for hundreds of samples. 
Tsinfer, ARG-Needle and Threads have 
been applied to datasets of hundreds of \emph{thousands} of samples.
With the explosive growth in dataset
sizes in recent years~\citep{caulfield2017national,turnbull2018100,
bycroft2018genome,backman2021exome,halldorsson2022sequences,uk2023whole,
all2024genomic,stark2024call,cook2025our}
there is a pressing need for more statistical
rigor in large scale ARG inference.
% This inability to connect the results of large-scale inference methods
% with a generative model
% is a fundamental limitation, and limits both the utility of 
% current point estimates as well as the possibilities for further
% inference. Specifically, we are unable to quantify uncertainty 
% around heuristic point estimates. We cannot explore ARG space
% around the point estimates using (e.g. MCMC), as we have no way definition
% of improvement. We cannot compare models of demography, or evaluate the 
% goodness of fit for model estimates.
A key problem facing the field is that the ARGs estimated by heuristic large-scale 
methods lack a concrete connection to population genetic theory because
we cannot compute their likelihood under the SMC or similar models. 
% The classical method~\citep{kuhner_maximum_2000} assumes complete knowledge about
% recombination events, and there is no clear way to compute a likelihood for 
% the approximate structures output by hueristic methods.

Computing the likelihood of an ARG under a model is a fundamental element of
probabilistic inference. The classical Kuhner, Yamato \& Felsenstein
\citeyearpar{kuhner_maximum_2000} formulation (henceforth: KYF)
works by assigning a probablility to each recombination and common
ancestor event, and to the inter-event  waiting times. The likelihood of an 
ARG is then a product over all these three classes of term.
% In essence, computing the KYF likelihood requires
% ``replaying'' the  state of Hudson's simulation
% algorithm~\citep{hudson_properties_1983,kelleher_efficient_2016} in order to
% compute the exact configuration of lineages and distribution
% of ancestral material at each time point~\citep{baumdicker_efficient_2021}.
% It therefore requires complete
% detail about each of these events to be stored. 
% Other approaches to defining the likelihood of an ARG under a coalescent-like 
% model follow the same basic event-by-event 
% strategy~\citep[e.g.][]{guo_recombination-aware_2022}.
% % [Are there others?  The bacterial stuff from Didelot etc seem to focus on
% % a tree-by-tree likelihood, rather than global graph, so not clear it's the 
% % same thing]
% Computing the likelihood under the SMC rather than the Hudson coalescent 
% simplifies bookkeeping somewhat,  % Dunno what I'm talking about here
% but a complete event-by-event description of the 
% input ARG is still required.
Although the KYF formulation is a natural and elegant way to describe 
the likelihood of an ARG, such complete detail is not present 
in many estimated ARGs~\citep{wong_general_2023} or even in most simulations.
Simulators such as the classical \texttt{ms}
program~\citep{hudson_generating_2002} output ARGs in a tree-by-tree format,
which does not contain sufficient detail for likelihood computation.
The \texttt{msprime}
simulator~\citep{kelleher_efficient_2016,baumdicker_efficient_2021} 
by default outputs a more complete and efficient representation of 
the simulated ancestry, but still does not provide the exhaustive
detail required.
Indeed, providing the information required to 
support likelihood calculations originally motivated the addition of
the \texttt{record\_full\_arg}
option to \texttt{msprime}~\citep{baumdicker_efficient_2021}.
Forwards-time simulations that output 
ARGs~\citep{kelleher_efficient_2018,haller_tree_2018} similarly do not
typically contain the level of detail required to compute a likelihood under
the KYF approach.

% [Things to include: locally unary nodes rather than explicit events.
% edge-centric formulation vs event centric. Graceful degradation of
% performance with incomplete ARGs vs complete inability to compute.]

In this paper we provide a new formulation of ARG likelihood which 
has much more lenient input requirements than KYF, and gracefully handles
incomplete and imprecise ARGs.
% In this paper we provide a necessary first step towards this goal
% of statistically rigorous large-scale ARG inference, by defining a function
% to compute the likelihood of a general class of ARG under the SMC.
To do this, we rederive the SMC under a backwards-time formulation,
modelling the process in terms of lineages and [something] rather
than events. Using this novel formulation, we then show how to 
easily compute the likelihood of any ARG under the SMC and outline 
an efficient algorithm to do so. We illustrate the flexibility of this
approach via [some examples] and conclude by discussing avenues
for future research opened up.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

% \section*{ARG likelihoods}

% In one of the first approaches to probilistic ARG inference that 
% regarded the ARG itself as an output rather than a latent variable,
% \cite{kuhner_maximum_2000} provided an explicit formula for 
% the likelihood of a realised ARG under the coalescent with
% recombination:
% \[
% \mathbb{P}[G|N, r] = \left(\frac{1}{2N}\right)^K
% r^X \exp\left(-\sum \left( \frac{1}{2N} \binom{k_i}{2} + r\ell_i\right)
% t_i\right),
% \]
% where [explain notation].
% Under this KYF formulation, the likelihood is computed by taking a 
% product of the probability of each realised event, and their 
% inter-event waiting times. While it is straightforward to write 
% this down, computation is not so easy, and in essence requires
% ``replaying'' the  state of Hudson's simulation
% algorithm~\citep{hudson_properties_1983,kelleher_efficient_2016} in order to
% compute the exact configuration of lineages and distribution
% of ancestral material at each time point~\citep{baumdicker_efficient_2021}.

% % Computing the likelihood of a realised ARG under the SMC simplifies the 
% % bookkeeping somewhat, because we no long have the possibility of 
% % ``trapped'' ancestral material, and each lineage can be represented 
% % by a single interval representing their ancestral material. Nonetheless,
% % we still require complete knowledge of the graph to compute the 
% % required probablilities.

% \section{Methods and Results}
% First, 
% we describe
% in more detail what sort of information we have in estimated ARGs.
% Then,
% we define the SMC process in a ``backwards-in-time'' way,
% rather than the usual ``along-the-genome'' way.
% Note that this is a
% reiteration of the original work of \citet{mcvean_approximating_2005}.

% In the subsequent section we connect the two and
% show how the ``backwards-in-time'' description
% makes it relatively straightforward to compute the likelihood
% for a more general class of estimated ARGs.
% Finally, we close with a demonstration that the likelihood computation can be used
% to do Bayesian inference on ancestral population sizes.

% % % % % % % % % % % % % % % % % %
\subsection*{ARGs} \label{par:recording}
Although the term Ancestral Recombination Graph (ARG) has been used to mean several things,
in the generic sense, it is a labeled graph structure that records
the inheritance of genetic material \citep{wong_general_2023}.
An ARG of the sort we consider describes the history of inheritance
of the genomic material of a focal set of individuals, or \emph{samples}.
Concretely, an ARG is equivalent to a collection of (non-contradictory) statements
of the form ``$c$ inherited from $p$ on the genomic segment from $x$ to $y$'',
where $c$ and $p$ are (haploid) genomes, and $x$ and $y$ are genetic coordinates.
We record each such statement in an ``edge'',
often summarized as the tuple $(p,c,x,y)$.
By ``describes the history'',
we mean that each node in the ARG represents a specific (ancestral) genome,
and that the ARG specifies the genealogical tree describing how the samples are related
at each point on the genome.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/smc_custom_2rows_area_full_hap.png}
\caption{ARG generated under the SMC, represented as a graph (top) and as a
series of local trees (bottom). To ensure the one-to-one correspondence
between both representations a node is recorded for each local tree whenever
a node is encountered along its ancestry path in the graph representation.
This results in nodes that (locally) have only a single child, and are therefore
(locally) unary. 
Nodes A, B and C are recombinants, i.e, have two parents in the graph
and are unary in the local trees. The blue shaded regions indicate
the range of possible times associated with these recombinants.
The blue arrows show the
standard left-to-right logic of the SMC whereby
the floating lineage coalesces randomly
with a the remaining portion of the local tree following recombination.
% By removing recombinant nodes, we limit the information on the time to each
% of these events to the windows (blue shaded array) delimited by the age of
% the child node and the minimum age of its parents.
}
\label{fig:smc-unary}
\end{figure}

An ARG can therefore be thought of either as a collection of inheritance
relationships between ancestral haplotypes,
or as a sequence of genealogical trees that have common node labels.
Fig.\ref{fig:smc-unary} shows a small ARG
that describes relationships between four samples on a 100bp piece of genome,
represented as both a graph (top) and a series of local trees (bottom).
Both structures represent the same information:
for instance, the fact that the ancestral genome at $A$
inherited the left half of its genome from ancestor $g$
and the right half from ancestor $f$
is seen in the fact that the first two trees have $A$ below $g$,
and the remaining have $A$ below $f$.
The blue dotted line from $A$ in the second tree to $f$ in the third tree
draws our attention to the fact that
the path traced by $A$'s ancestry switches at that point.
In fact, $A$ is in blue because it is a \emph{recombinant} node --
an ancestor that has a segment of genome inherited by a sample,
but that segment was inherited in two pieces from their parent.
Other nodes (labeled by black numbers) are either samples
or \emph{common ancestry} nodes --
ancestors that are the most recent common ancestor for two 
or more samples
at some place along the genome.

Crucial for the equivalence between the two representations is the presence of
nodes that only have a single child along one or more local trees --
i.e., are \emph{locally unary}~\citep{wong_general_2023}.
Recombinant nodes (in blue) have a single child in all trees.
However, common ancestry nodes can be locally unary as well:
for instance, sample $d$ inherited a longer stretch of genome from ancestor $f$
than sample $c$.
In the top plot, this is visible because the pink bar below node $f$
extends further to the left than does the blue bar,
while in the bottom plot, this is visible because node $f$ is parent to $d$
in the second tree, but not to node $c$, and so is unary.

%% Noved this section to section on smc
% A key observation of \citet{mcvean_approximating_2005}
% was that the sequence of trees along the genome
% generated by the SMC was Markovian \emph{after removing all unary nodes}.
% In other words, if we remove recombinant nodes entirely,
% and erase the unary portions of coalescent nodes,
% then the distribution of the next tree can be generated only knowing the current tree.
% (To see that the unary nodes makes the process non-Markov,
% consider that a node in a tree might be unary either because it's about to be a coalescent node
% or it already has been:
% looking at the second tree (Fig.\ref{fig:smc-unary}), we know node 5 will be coalescent further down the sequence
% because it was not present in the first tree.
% So, the first tree gives us information about subsequent trees
% that the second tree alone could not.)
% However, removing these unary portions of nodes seems odd
% when one considers the consequences of the inheritance paths through the
% ARG in the top figure.
% As pictured, both figures tell us that
% 3 inherits the middle bit of the genome from 7 through 5.
% However, if we erase 5 from the second tree,
% does this suggest that 3 inherited a single segment from 7 along two distinct paths?
% The graph-view of the ARG (above) makes it clear that this is unlikely:
% it is in fact reasonable to leave in these unary sections.

These locally unary nodes enable
us to uniquely identify all lineages that were hit by a recombination event,
even if we simplify 
the ARG by removing all recombinant nodes
($A$, $B$, and $C$ in Fig.\ref{fig:smc-unary}).
In the absence of these recombinant nodes, a recombination can be observed
easily enough as a change in parent going from one genomic region to the next.
Without the recombinant nodes we no longer know the precise time of each recombination event,
but we do still have constraints on it;
for instance, we know that $C$ happened somewhere between node $f$ and node $h$.
More generally, a recombinant node must happen somewhere between the child node
and the most recent of the parent nodes.
This is again easier to visualise using the
graph representation (blue shaded area, Fig.\ref{fig:smc-unary}).
Although these unary nodes are useful and to some degree clearly inferrable,
they are not always represented by simulators or ARG inference methods.
However, to our understanding this is because
the field is used to thinking in terms of the trees
rather than in terms of relationships between haplotypes,
not because of any question of knowability.
% JK I think we can but this out without-loss now given Halley's paper?
% Currently,
% \argweaver{} \citep{rasmussen_genome-wide_2014},
% \tsinfer{} \citep{kelleher_inferring_2019},
% \kwarg{} \citep{ignatieva_kwarg_2021}, and
% \texttt{SINGER} \citep{deng_robust_2024} estimate ARGs 
% including unary portions of coalescence nodes.
% However, it is unclear to what degree these are well-inferred,
% as most evaluation of inferred ARGs thus far has focused on marginal trees
% (e.g., \citet{brandt2022evaluation, kelleher_inferring_2019},
% but see \citet{deng2021distribution})
% and thus completely ignore such unary nodes
% (and their implications for inherited haplotypes).
See~\cite{fritze2024forest} for more discussion on unary nodes
and haplotype aware ARG comparison metrics.

% % % % % % % % % % % % % % % % % %
\subsection*{SMC backwards-in-time}\label{par:description}

% coalescent introduction: backwards-in-time
% left-to-right: Wiuf-Hein (difficult)
% SMC: although easy to reason about left-to-right sense
% 
The coalescent process describes the ancestry of a set of sampled genomes.
The original algorithm 
to simulate the coalescent with
recombination was formulated backwards in time
by Hudson \citeyearpar{hudson_properties_1983}. 
Wiuf and Hein later described
a tree-by-tree method to simulate the same stochastic process 
\citep{wiuf_recombination_1999}.
This along-the-genome approach is considerably more complex as the next 
tree to be simulated depends on all previous trees.
This idea however provided the basis of the 
SMC \citep{mcvean_approximating_2005}, restricting the set of all state
space transitions possible under the Hudson coalescent to
obtain a process that is Markovian both backwards-in-time, 
as well as along-the-genome.
In its backwards-in-time formulation, the SMC only requires a
simple modification to the coalescent with recombination.
Instead of allowing any pair of lineages to coalesce,
the SMC is the process in which only lineages with \emph{overlapping} 
genomic intervals may coalesce. The SMC' is a refinement 
of the SMC which also allows lineages with abutting genomic intervals to 
coalesce~\citep{marjoram_fast_2006}.
Although \citet{mcvean_approximating_2005} verbally described the SMC
using this backwards-in-time description, their analysis
was almost entirely in left-to-right terms,
and subsequent papers followed this trend
\citep{li_inference_2011,paul_accurate_2011,schiffels_inferring_2014,
rasmussen_genome-wide_2014}.

Linking the resulting tree-valued process along the genome back to
the relationship between the ``graph'' and
``local trees'' representation of ARGs, 
the Markovian property can be formulated as follows.
The distribution of the next tree within the series
of local trees can be generated only knowing the current tree,
provided that we remove recombinant nodes entirely,
and erase the unary portions of coalescent nodes.
To see that the locally unary nodes make the process non-Markov,
consider that a node in a tree might be unary either because
it's about to be (in the left-to-right sense) a coalescent node 
or it already has been:
looking at the second tree in Fig.\ref{fig:smc-unary},
we know node $f$ will be coalescent further along the sequence
because it was not present in the first tree.
So, the first tree gives us information about subsequent trees
that the second tree alone could not.
Further note that by limiting common ancestry events to overlapping 
pairs of lineages, we are guaranteed that all nodes in the ARG 
only carry ancestral material, that is,
genetic material that is inherited by either one of a set of samples
\citep{wiuf_ancestry_1999}.
We therefore require a single half-open interval to describe
the ancestral
material associated with each node in the ARG.

We now define the process more formally, by reintroducing
the notation of \citet{mcvean_approximating_2005} required to describe
the likelihood computation. This section
does not deviate from the original description of the SMC apart from
how coalescable pairs are counted. This subtle nuance
greatly simplifies the ability to keep track of coalescable pairs going
backwards in time.
At any point in time, the state of the process is 
the set of labeled lineages extant at time $t$,
$L(t) = \{X_j(t)\}$.
A lineage $X_j(t) = [x_{j}, y_{j})$ is labeled by 
an integer $j$ and defines the 
half-open genomic interval of its ancestral material.
% Ea
% where each lineage $X_j(t)$ is labeled by an integer $j$
% and represented by a 
% ancestral material. 
% Any lineage can thus be summarized as 
% for some interval endpoints $x_{j} < y_{j}$.
Because we look backwards in time, $t=0$ is today, and 
$L(0)$ consists of $n$ sampled lineages, labelled $0$ to $n-1$,
each represented by a single interval spanning the entire genome.
Then, the process evolves by a succession of common ancestor and recombination
events until each segment of ancestral material is only present in one lineage.
The waiting time to next event is determined by these two competing processes
with exponentially distributed waiting times as outlined below.

Recombination is described by a Poisson process of rate $r$ per unit 
of genomic length and time:
so, if $T(t) = \sum_j |X_j(t)|$ is total length of ancestral material
carried by lineages at time $t$,
then recombination occurs at instantaneous rate $rT(t)$.
A recombination to the left of $x$
(i.e., between base pairs $x$ and $x-1$)
with $x_{j} < x <y_{j}$ splits lineage $j$ into two new lineages,
given new, unique labels. We then remove lineage $j$
and add the two new lineages to the state $L(t)$.
This operation keeps the total amount of ancestral material unchanged.

Assuming a randomly mating, diploid population of constant size $N_e$,
common ancestry occurs between
two \emph{overlapping} lineages at rate $\lambda = 1/(2N_e)$.
Here we deviate from the McVean and Cardin formulation,
and define a strict total order on all lineages in $L(t)$:
for each lineage $X_j(t) = [x_{j}, y_{j})$ we define its 
order by the left coordinate $x_j$ and label $j$ (to break ties).
The instantaneous rate of coalescence then equals $\lambda$ 
multiplied by the number of overlapping pairs,
i.e., $\lambda \sum_{i \neq j} I_{jk}$,
where
\begin{equation} \label{def:coal}
I_{jk} = \begin{cases}
    1 & X_j > X_k \text{ and } y_k > x_j \\
    0 & \text{otherwise.}
\end{cases}
\end{equation}
We say that lineage $k$ can coalesce \emph{into} lineage $j$ if $I_{jk} = 1$.
From the definition of the strict total order it follows that if $I_{jk} = 1$
then $I_{kj} = 0$. 
The newly formed lineage acquires the union of both intervals,
the original lineages are removed, and $L(t)$ is updated accordingly.
% [JK notes: need to explain here that this deliberately introduced 
% asymmetry doesn't affect the actual process. Also need to comment on 
% how the process terminates. Basically has the same subtleties as McVean and 
% Cardin, and not important for our purposes.]
For later use, also define $C_j(t)$ to be the set of
lineages that can coalesce into lineage $j$, i.e., 
$C_j(t) = \{X_k \in L(t) | I_{jk}(t) = 1\}$; 
thus, $|C_j(t)| = I_{j}(t) = \sum_{k} I_{jk}(t)$,
and the total rate of coalescence at time $t$ is $\sum_{j} |C_j(t)|$. 
Because $I_{jk}$ is defined in terms of a total order, the sets $C_j(t)$ are disjoint,
which will simplify likelihood computations later. In particular, although
recombination affects \emph{which} lineages can coalesce into $j$, it does not
change the \emph{number} of such lineages: in other words, a recombination
event changes $C_j(t)$ but not $I_{j}(t)$.

%% I have updated the description of the process below to avoid having to
% define yet another symbol.
%Another useful quantity to know is the number of lineages carrying material
%at each point on the genome:
%for a location $z$ and time $t$, this is
%\begin{equation}
%    N(t,z) = \vert\{j : z \in X_j(t) \} \vert
%\end{equation}

% [Do we need to get into this above-the-root subtelty here?]
% To ensure that the ancestral material carried by a single lineage can at
% all time be represented by a single half-open interval, we only update the
% endpoints of the involved lineages following common ancestry or recombination events
% such that only the minimal half-open interval whose endpoints have not fully coalesced
% is used to update $L(t)$. Note that empty intervals are not added to the state.
% %By reasoning about the process in this way we can keep representing 
% % all lineages by means of a single half-open interval until $\vert L(t) \vert = 0$
% % and the process terminates.

% %After any coalescence or recombination event,
% %the segment(s) $[x_j,y_j)$ (and $[x_k, y_k]$) carried by the involved lineage(s)
% %are replaced by a single segment $[x, y)$ whose endpoints have not fully
% %coalesced, i.e., $x = \min\{ z \ge \min(x_j, x_k) : N(t,z) > 1\}$,
% %and $y = \sup\{ z \le max(y_j, y_k) : N(t,z) > 1\}$. Note that empty intervals
% %are not added back to $L(t)$.

% Finally, by describing the SMC
% backwards in time, some intervals contained within the half-open
% interval representing a lineage might have fully coalesced before all the genetic
% material carried by that lineage has. As noted by \citet{mcvean_approximating_2005},
% such intervals can be observed as the stalks on the marginal genealogies (e.g.\ marginal trees
% for $[25, 81)$ Fig. \ref{fig:smc-unary}). Recombination events occurring within these smaller
% intervals however do not affect the marginal genealogies.

%The process of recombination and common ancestry events continues 
%Finally, there is one more operation that allows the process to terminate:
%portions of the edges of lineages that have entirely coalesced
%(i.e., that are ancestral to all samples) are removed.

%Formally, after any coalescence or recombination event,
%the segment $[x,y)$ carried by the lineage
%is replaced by the largest (pair of) segment(s) whose endpoints have not
%coalesced, i.e., $x$ is replaced by $\min\{ z \ge x : N(t,z) > 1\}$,
%and $y$ is replaced by $\sup\{ z \le y : N(t,z) > 1\}$.
%By reasoning about the process in this way we can keep representing 
%all lineages by means of a single half-open interval throughout.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{ARG likelihood} \label{par:liks}
We have defined the SMC process in terms of rates of various types of events.
In general, the likelihood for a continuous-time Markov process specified in this way
is $\exp(-\Lambda) \prod_i \lambda_i$,
where $\lambda_i$ is the rate of the $i^\text{th}$ realized event,
and $\Lambda$ is the sum of all rates of all possible events
integrated over the entire process
(also called the ``total hazard''). 
As discussed in the Introduction, in many cases these events are not what is 
estimated by inference methods, which produce, more fundamentally, a collection
of genetic inheritance relationships (edges) between ancestral haplotypes (nodes).
Our goal, therefore, is to decompose the overall likelihood into the
per-node and edge contributions, allowing us to compute likelihoods
for the incomplete ancestries estimated by current methods.
% , as well as graceful degradation in the presence of incomplete information.

%\comment{if we want to spell it out even further:
%Note that the time to the next event is distributed as $\Lambda \exp(-\Lambda)$. And
%the probablility that the next event is of type $i$ is $\lambda_i / \Lambda$.}

%where each edge $e$ is of the form $e = (p,c,x,y)$.
%Recall that this means that node $c$ inherited genetic material from node $p$
%over a region with span from $x$ to $y$ along the genome
%(see Figure~\ref{fig:likelihood}).
%So, $x$ and $y$ are genomic coordinates,
%while $c$ and $p$ are in $N$ (i.e., distinct sampled or ancestral haplotypes).

%Because the process of recombination only affects one lineage at a time
%(and does not even depend on the state of other lineages).
%On the other hand, common  events necessarily involve more than one lineage.
%However, the asymmetric way we define common ancestry events above
%allows us to uniquely associate each event with a single lineage,
%which allows us to cleanly factor the likelihood across edges and nodes.
% latter realized events, factor hazard across the edges.

% The structure of the ARG, as a collection of \emph{edges} $E$ and \emph{nodes} $N$,
% as well as the definition of both recombination and common ancestry events, allows us
% to cleanly factor the likelihood of the ARG across its edges and nodes.
% Not only does recombination
% naturally only affect one lineage at a time without even depending
% on the state of other lineages. Common ancestry events can be equally
% associated with a single edge given definition \eqref{def:coal}.
% The next paragraphs detail precisely how the number and the nature of the realized events
% can be determined based on the number of
% parents and children of each of the nodes in $N$ as well as how the total hazard can be
% factorized into the individual contributions of each of the edges.

Let $\deg_o(u) = \vert\{e \in E : p_e=u\}\vert$ be the out-degree of a node $u$,
i.e., the number of edges having $u$ as a parent.
Because mergers are binary in the SMC,
this must have been the result of $\deg(u) - 1$ common ancestry events,
and so this contributes a factor of $\lambda^{\deg_o(u)-1}$ to the likelihood
for nodes with any children.
Note however that this description is general and holds for the case of
more-than-binary mergers. By treating such polytomies as if there were zero-length edges
between a sequence of mergers, the order of these unresolved mergers does 
not affect the likelihood. Under the SMC the likelihood of a zero-length edge is zero.
This is not special: since edge lengths are continuously distributed,
the likelihood of any particular value is also zero.

Similarly,
let $\deg_i(u) = \vert\{e \in E : c_e=u\}\vert$ be the in-degree of a node $u$,
i.e., the number of edges having $u$ as a child.
Each additional edge that is ancestral to a given node
implies one additional ancestral recombination event,
and thus a factor of  $r^{\deg_i(u)-1}$ in the likelihood,
if the node has any parents.
Put together, these two contribute
\begin{equation}\label{eq:coal}
    \prod_{u \in N} \lambda^{\deg_o(u)-1} r^{\deg_i(u)-1}  
    =
    r^{|E|-|N|+n_r} 
    \lambda^{|E|-|N|+n_s} 
\end{equation}
to the likelihood,
where $n_r$ is the number of nodes with in-degree 0 (roots),
and $n_s$ is the number of nodes with out-degree 0 (usually, samples).
(The simple form follows because each edge contributes exactly one
to some node's out-degree and some node's in-degree.)

\begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{figures/single_event.png}
    \caption{
    Graphical representation of the events associated with the edge $(p,c,x,y)$.
    The lineage carried by $c$ (dark green bar) spans $[x_c,y)$ 
    until a recombination event (blue
    dotted line) splits it into two segments. Left and right hand segments coalesce
    (with the pink and yellow lineages, respectively)
    at times $t_{p'}$ and $t_p$ respectively, bounding the time to recombination by
    % JK used to be: $(c. \min(t_p, t_{p'})$
    $\min(t_p, t_{p'}$.
    % JK: not sure this helps? 
    % Note that the contribution to the recombination hazard
    % by edge $(p,c,x,y)$ only covers ``half'' of the green shaded area. The other
    % half is contributed by edge $(p', c, x_c, x)$. 
    \label{fig:likelihood}
    }
\end{figure}

Now consider the ``hazard'' associated with recombination.
Begin with an edge $e = (p,c,x,y)$.
This implies that there was a sequence of lineages
from the time of the child $c$ (call this time $t_c$)
back to the time of the parent $p$ (again, $t_p$),
that contained the segment $[x,y)$,
and the lineage that directly inherited from $p$
had the segment $[x,y)$ (see Fig.~\ref{fig:likelihood}).
We therefore know there was no recombination between $x$ and $y$
on any of those lineages 
over this time span,
the probability of which is
$\exp(-r \times \mathcal{A})$, where $r$ is the recombination rate
and $\mathcal{A}$ is the total area (eligible links times depth) of the edge.
So, the first contribution to the likelihood from this edge $e$ is
\begin{equation}\label{eq:span}
A_e(r) = \exp(-r (y-x - \mathbbm{1}_{x > x_c})(t_p - t_{c})) .
\end{equation}
Note that if a recombination did occur, then the number of links not
affected by recombination is 1 less (see eq.~\eqref{eq:depth}).

Finally, we turn to the hazard associated with common ancestry events.
To do this,
we need only consider what happens along the left side of each edge.
Recall that we define a total order on edges,
by ordering by left endpoint and breaking ties by child
(i.e., $(p,c,x,y) < (p',c',x',y')$
if $x<x'$, or if $x=x'$ and $c<c'$).
The instantaneous rate of coalescence of a lineage whose left endpoint is at $x$
at a given time is equal to $\lambda$ multiplied
by the number of earlier lineages that overlap $x$ at that time.
Using this total order, we can define
$I_e(t)$ to be the total number of earlier edges that overlap edge $e$
and are present at time $t$.
Then, the cumulative hazard for common ancestry events of edge $e$
between time $s$ and $t$ is given by $F(e, s, t) = \int_{s}^{t} I_{e}(z) dz$.
If the edge $e = (p,c,x,y)$
is the leftmost segment of material inherited by node $c$
(e.g., the left hand segment ancestral to node $c$
in Fig.~\ref{fig:likelihood}),
then no recombination occurred along this edge and 
this edge simply contributes $\exp(-F(e,t_c,t_p))$ to the likelihood.
However, if the edge is \emph{not} the leftmost segment
(as shown in Fig.~\ref{fig:likelihood}),
then it was initiated by a new recombination breakpoint,
and we need to integrate over possible times the breakpoint occurred.
If this is the case, there is another edge $(p',c,x',x)$ with the same child node,
immediately adjacent to the focal edge $(p,c,x,y)$.
The recombination event that split the two must have happened between $t_c$
and the smaller of $t_p$ and $t_{p'}$. Note that two events must have happened
between $t_c$ and $\min(t_p, t_{p'})$; taking all this into consideration,
the contribution to the likelihood here is
\begin{equation}\label{eq:depth}
B_e(r, \lambda) = \begin{cases}
    e^{-\lambda F(e, t_c, t_p)}
        & \qquad \text{if } x=x_{c} \\
    \int_{t_c}^{t_{p} \wedge t_{p'}} e^{- r s} e^{-\lambda F(e, s, t_{p})} ds
        & \qquad \text{if } x>x_{c} ,
\end{cases}
\end{equation}
where $x_c$ is the leftmost position of the edges whose children are $c$.
% In that case $x=x_{c0}$ and $c$ did not coalesce until $t_p$.
% Alternatively, $x$ can be a new recombination break point that
% occurred on the lineage represented by $c$ at some time $s$ between $t_{c}$
% and $\min(t_p, t_{p^{\prime}})$, where $p^{\prime}$ is the parent of the
% edge $(c, p^{\prime}, X_{cp^{\prime}})$ ending at $x$. The new lineage formed
% by the recombination event and starting at $x$ can be given a temporary label
% $c_s$. In both cases, the cumulative hazard of a lineage $a$ coalescing
% between time $s$ and $t$ is given by $F(c, s, t) = \int_{s}^{t} I_{c}(u)du$
% such that:
%\comment{This is the old equation; I think the $e^{-rs}$ should not be there,
%    since it's already taken into account by \eqref{eq:span}.
%    Plus, the $\lambda$ and $r$ factors are now elsewhere.}
%\begin{equation}\label{eq:depth_old}
%B_{(c, p)}(\theta) = \begin{cases}
%e^{-\lambda F(c, t_c, t_p)} \lambda^{I_{cd}(t_p)} & x=x_{c0} \\
%\int_{t_c}^{t_{p} \wedge t_{p^{\prime}}} r e^{-rs} e^{-\lambda F(c_s, s, t_{p})} ds \lambda^{I_{c_{s}d}(t_p)} & x=x_{c_{s}0}>x_{c0} \\
%\end{cases}
%\end{equation}
%\comment{Do we still need this bit?}
%Here, the (second) exponential term gives the probability of not observing a
%coalescence event before $t_p$. In case of a recombination, $re^{-rs}$ further
%gives the probability of observing a recombination
%at time $s$ given a Poisson process along position $x$ with rate $r$.
%The time of the recombination is integrated out.
%The last term gives the point probability density of the coalescence event
%that terminates the edge. The indicator function is used to resolve
%the non-reciprocal nature of coalescence events as defined in \ref{def:coal}.
The overall likelihood of an ARG defined by the set of edges $E$ and nodes $N$
given parameters $\theta$ is then given by
\begin{equation}\label{eq:full-lik}
    \mathcal{L}(\mathcal{G}|r, \lambda)
    =
    (r)^{|E|-|N|+n_r} (\lambda)^{|E|-|N|+n_s} \prod_{(p, c) \in E} A_e(r) B_e(r, \lambda) .
\end{equation}
One significant advantage of this formulation is that the computation time
is linear in the number of edges, and an efficient algorithm exists
to keep track of $I_e(t)$ (see Appendix).
% Although the first two factors are easily expressed in terms of numbers of edges and nodes,
% we will see below
% that in practice it is convenient to absorb these terms into per-edge contributions.
% We can also accommodate stepwise changes in the 
% coalescence rate (or recombination rate)
% through time by intersecting the internode intervals with the rate change
% intervals; rate changes along the genome can be incorporated similarly.
Note that we have only considered the topology and branch length information here
an not considered mutational information.
The likelihood of a set of mutations given an ARG is
\begin{equation}\label{eq:lik-mut}
    \mathcal{L}(\mathcal{D}|\mathcal{G}, \mu)
    =
    \frac{1}{\mathcal{M}!} \prod_{(p, c, x, y) \in E} e^{-(y-x)(t_p-t_c)\mu^{|m_e|}}
\end{equation}
where $\mathcal{M}$ the total number of mutations and $m_e$
the number of mutations along edge $(c,p)$ \citep{mahmoudi_bayesian_2022}.

% % % % % % % % % % % % % % % % % %
\subsection*{Illustrative examples}
% likelihood surface
% The last paragraph aims to highlight the flexibility of this approach 
% and the impact of the various
% degrees of completeness of the ARG on the likelihood.
% Note that the figures in this paragraph only serve this exact purpose,
% we are not claiming to have designed a new parameter estimation approach.

Eq.~\eqref{eq:full-lik} allows us to compute a likelihood for \emph{any}
collection of nodes (ancestral haplotypes that lived at some time 
$t$ in the past) and edges (a record that child node $c$ 
inherited the genomic interval $[x, y)$ from parent node $p$)
under the SMC. Other than some basic consistency constraints (e.g.\ parents
must be older than children) there are no requirements on the topology
described, and in particular, no requirement that the events from the 
model (the SMC) should be explicitly identified in the graph. Our 
formulation reasons about the events that \emph{could have resulted in}
the observed nodes and edges, and therefore frees us from the requirement
of having to explicitly write those down. This gives a great deal 
of flexibility, which we illustrate in this section via some examples.

The first example concerns the simplest possible polytomy (see
the Discussion for why we consider polytomies important).
Suppose we have a graph consisting of four nodes, $c_1,c_2,c_3,p$ and three edges
${(p, c_j, 0, 1): 1 \leq j \leq 3}$. The child nodes are samples at time
zero, and the parent has time $1$. The likelihood of this ARG is undefined
under HKY because the three-way star topology described cannot be interpreted as 
a sequence of binary coalescence events. However, it is clearly 
\emph{consistent with} such a sequence of events, which we can reason
about probabilistically. In our formulation we [Peter/Gertjan, help me 
out any hold my hand a bit here. How exactly to we assign a likelihood
to such a polytomy? Why do I get a larger likelihood for the ternary
tree described above than an example in which I break the polytomy? 
Naively, I would have thought that the likelihood should have been 
lower, but I clearly don't understand likelihoods...]

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/likratio.png}
    \caption{TODO add A/B and change ``unary regions''.
    Log likelihood-ratio curves for ARGs simulated under the SMC with
    various degrees of completeness. The likelihood ratio is computed with
    respect to the true recombination rate. In case of the full ARG the likelihood
    under the SMC has been computed using the exact time to the recombination event.
    Simulation performed with \texttt{msprime}, parameters: $L=1$Mb, $20$ diploid samples,
    $r=2.5e-8$, $N_e=1e4$.}
    \label{fig:lik-surface}
\end{figure}

Our second example shows a different aspect of this robustness to missing 
data. In Fig~\ref{fig:lik-surface}A we can see the likelihood-ratio curves
obtained when we compute the likelihood of an example simulation,
in which the ARG is at three different levels of completeness
(illustrated in Fig~\ref{fig:lik-surface}B).
We first run a simulation under the SMC
using the \texttt{record\_full\_arg} option in \texttt{msprime}. This 
ARG explicitly identifies all of the recombination and common 
ancestor events that occurred in the simulation, and 
we can therefore compute a likelihood under the HKY formulation
(for example, using the \texttt{msprime.log\_arg\_likelihood} 
function). The example full ARG in 
Fig~\ref{fig:lik-surface}B has three nodes that are ``locally
unary''~\citep{wong_general_2023}. Nodes 6 and 7 correspond to the 
left and right parents of the recombination event causing the 
breakpoint at position 536~\cite[we need two nodes to uniquely identify
 the passage of ancestral material through recombination 
events; see][]{baumdicker_efficient_2021}, and therefore present
in only those trees. 
Node 8 is a coalescent node in the right-hand tree, but unary
in the left-hand tree, marking the passage of the lineage through
this ancestral haplotype (similar to node g in Fig~\ref{fig:smc-unary}).
Fig~\ref{fig:lik-surface}A shows that, remarkably, when we 
remove the recombination nodes (e.g., 6/7) but keep the locally unary nodes (e.g., 8)
and compute the likelihood using eq.~\eqref{eq:full-lik},
there is very little difference in the likelihood ratio
curve from the exact value with complete input. 
Most importantly, the minimum occurs at the true value of the 
recombination parameter.
When we further simplify
the ARG to remove all unary nodes~\citep{wong_general_2023,kelleher_efficient_2018} 
we obtain an ARG in the standard format used by \texttt{msprime}.
Although there is substantial information about the correlation
structure of trees in this representation, we lose important
information about recombination. This is demonstrated in 
Fig~\ref{fig:lik-surface}A, where the corresponding likelihood ratio curve
is clearly biased.

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/3-slices-whitespace.png}
    \caption{Simulated ARG under Hudson coalescent (left). $N_e$ is constant
    for each of the 3 time slices: $[(0:1000): 10 000, (1000:5000): 5000,
    (5000:\infty):10 000]$. Remaining parameters: $L=10$kb, $20$ diploid samples,
    $r=2.5e-8$, $148$ edges in ARG. A single $N_e$ value was estimated for each
    slice of the ARG as defined by the true population size change time points.
    The inferred posterior distribution (right) using the SMC-likelihood was
    obtained using a uniform prior. Number of edges per time slice
    (from most recent slice to oldest): $[68, 57, 23]$. Locally unary nodes were
    omitted for clarity.}
    \label{fig:3-arg-slices}
\end{figure}

Our final example concerns [TODO motivate this example; why is 
this interesting/useful? Presumably we don't have to jointly estimate
 the ARG with demography, but can do so post-hoc? But, couldn't 
you do that anyway if you had a full ARG using HKY?]
% slicing of the ARG
By explicitly associating a single likelihood with each edge,
we can compute the likelihood for any slice of the ARG taken in
time and/or along the genome.
In the case of a fully precise ARG, with a known time to recombination,
this is true as long as the slicing operation does not erase the
information on whether an edge is associated with the left or
right hand segment generated by a recombination event. In the absence
of the exact time to recombination, the integral in
eq.~\eqref{eq:depth} means that the likelihood
for the entire ARG does not factorize into the likelihoods of these
individual slices. More precisely, the likelihood associated with any
edge that is initiated by a recombination breakpoint cannot be factorized
into the product of two integrals of the same form.
However, as long as the individual slices of the ARG contain sufficient
information we can approximate the true integral as the product of the
integral of the slices. Fig.~\ref{fig:3-arg-slices} demonstrates this idea
on a toy example. We simulated an ARG under human-like parameters with
two effective population size changes, and subsequently computed the posterior
distribution on the three effective population sizes
using a uniform prior on each slice independently.
The result is reasonably accurate, despite only using 10kb of sequence.
This approach therefore strengthens
the inherent ability of ARGs to provide us with temporal
resolution and context to a dataset of genetic variation.
It also opens up the opportunity for more efficient optimization
strategies during inference given that each slice of the ARG can be
processed independently.

% And it further highlights the flexibility
% of the algorithm. Simply stated and from a data structure perspective,
% we can compute a likelihood under the SMC for any valid tree sequence.
% The \texttt{tskit} library provides a wide range
% of subsetting operations on tree sequences allowing us to slice
% the ARG in time and/or space, as well as to reduce the number of
% tracked samples. Any such subset of the tree sequence is itself
% a valid tree sequence for which we can again compute a likelihood.

\section*{Discussion}

The key contribution of this paper is to rederive the classical KYF
ARG likelihood function in a way that supports partial and imprecise
input, using a novel backwards-time derivation of the SMC.
The underlying process that we reason about probabilistically is still
composed of events affecting lineages, but, acknowledging that we 
do not necessarily observe or estimate these events, we calculate 
the possible contributions of various type of event to the inheritance
relationships that are provided as input.
Although we can associate a likelihood with any ARG in \texttt{tskit}
format, there are some caveats. By treating each edge independently we can
naturally deal with the various degrees of completeness to which inheritance
is documented by the different ARG inference methods: e.g., polytomies, absence
of locally unary coalescence nodes, unpreserved identify of internal nodes
across local trees, etc.
However, many of these imprecisions violate the implicit
assumptions of the likelihood computation.
For example, we assume the 
presence of locally unary nodes to correctly identify the
full extent of all lineages going back through time and therefore recombination events.
Without these unary nodes, our algorithm will detect a recombination
event along both coalescing lineages in case of a (sub)tree height changes along
the ARG as both edges would switch parent nodes. 
Locally unary nodes are not a common aspect of current ARG research, but 
are estimated by \texttt{tsinfer}, and can be imputed to high
degree of accuracy from fully simplified ARGs~\citep{fritze2024forest}.

[A paragraph about possililties for inference]

[A paragraph about why what we've presented is useful, and a good starting
point for other research. We provide an efficient algorithm and 
reference implementation]

% This can be mitigated by limiting
% the number of `inferrable' recombination events per coalescence event to 1
% (see fig \ref{sup:fig:rec-correction}).
% \comment{I'm not editing this in since I'm not sure I've got it right:
% I think that as written, our algorithm would count an extra recombination
% and and extra coalescence for any edge that got split up by removing the unary bits.
% This could be corrected for in the algorithm, but it would require extra bookkeeping,
% and make the likelihood not so easily a product over edges. (?)}
% 
% \comment{this was previously part of the discussion, have grouped it here.}
% Although their presence is crucial to our ability to embed any local tree in the
% ARG, the concept and their importance are still relatively new (but see
% \citet{wong_general_2023}).
% We have briefly touched upon the observation that in
% the presence of unary nodes we seem to augment the left-to-right logic of the SMC
% with some additional information. Yet, we believe that the topic of locally
% unary nodes in general requires further exploration.
% And more specifically, we see it as future work to fully quantify to what
% extent the absence of these locally unary nodes will affect the shape of the
% likelihood surface given any set of parameters.

%%% remarks on precision of information + useage of likelihood to either
%%% estimate parameters (not goal here) or use as instrument to improve accuracy to infer ARGs

% [JK: this needs significant work]
% Here, we have introduced a scalable algorithm to compute likelihoods
% under the SMC for a general class of ARGs. By formulating the SMC backwards in time
% we were able to associate a single likelihood value with each edge in the ARG.
% The compatibility of the algorithm with the \texttt{tskit} library allows us to
% exploit its associated efficiencies both for the current work and its future
% applications. This work will contribute significantly
% to ARG-based analysis becoming a standard part of the population
% and statistical genetics toolkit.
% In particular, we see two main applications for the work presented here.
% Firstly, the ability to associate inferred ARGs with a likelihood given a demographic
% model and an associated set of parameters, will help bridge the gap between the tremendous recent
% advances in ARG inference and the current state of the art statistical inference of
% past demography.

% The second and probably the most important application of the ability to associate a more
% general class of ARGs to a principled model is the ability to boost the current efforts
% in ARG inference. 
% % great starting point for MCMC
% Any inferred ARG can be used a good starting point
% for a future MCMC-sampler. This potentially implies a massive runtime reduction by
% reducing the necessary burn-in period. Furthermore, by integrating out the exact time
% to the recombination event we further restrict the size of the ARG space that needs to
% be explored. This MCMC-sampler would again rely on the tree sequence data structure to
% efficiently generate and evaluate any new proposal (see \citep{mahmoudi_bayesian_2022}).
% % hybrid ARG-inference: scalability + bringing ARG down from pedigree dominated phase
% To further improve scalability, hybrid ARG-inference methods could be designed which
% rely on heuristic approaches to bring the ARG down from the initial pedigree-dominated phase,
% and subsequently use a principled approach to perform ARG inference for the second phase.
% This idea resembles the tried and tested approach of hybrid backwards-in-time simulations 
% that combine Wright-Fisher dynamics in the very recent past and coalescent simulations 
% for the remaining time \citep{bhaskar_distortion_2014, nelson_accounting_2020}. This hybrid
% approach has enabled large scale simulations while avoiding the documented biases of the
% coalescent relative to the Wright-Fisher model \citep{bhaskar_distortion_2014, wakeley_gene_2012}. 
% The same approach could be thus used for ARG-inference where one can rely on a heuristic 
% inference method for the recent past, while using a model-based approach beyond that cutoff point.

% Finally, acknowledging the uncertainty associated with ARG inference does not
% necessarily require a full-blown MCMC sampler. Instead, recognising those parts of
% the ARG for which we lack information for confident reconstruction and resolving
% them stochastically by sampling from the coalescent rather than completely arbitrarily
% would already be a major step forward.

% % this part is only a very rough idea, might be better to leave it out if we can't make it more concrete.
% We believe that quantifying uncertainty does not have to be associated with sampling
% from the space of all compatible genealogies. Uncertainty could also
% be quantified and passed on as metadata to the nodes in the data format storing the ARG.

% [Note: a point to make here is that the idea of combining hueristic large-scale 
% inference with model based inference is analogous to the combination of 
% coalescent simulation and forwards time simulation. We can also draw lessons
% from developers of different methods cooperating and using the tskit library
% to enable this.]

% Additionally, there is a rich suite of tools growing around
% the \texttt{tskit} ARG library~\citep{ralph_efficiently_2020}
% including simulation~\citep{kelleher_efficient_2016,kelleher_efficient_2018,
% adrion_community_2020,terasaki_geonomics_2021,
% baumdicker_efficient_2021,
% korfmann_weak_2022, petr_slendr_2023,tagami_tstrait_2024},
% ARG inference~\citep{kelleher_inferring_2019,speidel_method_2019,
% wohns_unified_2022,mahmoudi_bayesian_2022,
% rasmussen_espalier_2022,
% zhang_biobank-scale_2023,zhan_towards_2023} and analysis
% packages
% % etc
% \citep{nowbandegani_extremely_2023,fan_likelihood_2023}.
%% NOTE: this is some orphan text from the intro that can be rejiggered for the 
% discussion. We can talk about precision of recombination estimates a bit 
% here I think, as it is an important topic.

% Both approaches have further formed the basis (of many) of the recent advancements in
% ARG inference methods \citep{rasmussen_genome-wide_2014, heine_bridging_2018,
% kelleher_inferring_2019, speidel_method_2019, rasmussen_espalier_2022, zhang_biobank-scale_2023}.
% A further classification of these methods could be made based on the level of detail
% that is provided on recombination events. The first real breakthrough in large-scale
% ARG inference, \argweaver, heads the first group.
% The ARGs sampled by \argweaver from the posterior distribution
% given the observed variational data, are all fully precise and detailed in the sense
% that all events are explicitly inferred. Note however that the set of possible node times
% has to be limited to enable the usage of an HMM. The other more recent methods, although
% still SMC-based, have added heuristic pre- and/or post-processing steps
% that have allowed them to extend their approach to larger sample sizes. We should also
% mention \kwarg and \texttt{ARGinfer} in this category of methods producing fully detailed ARGs
% although these are not based on the SMC or the copying algorithm. In fact they represent
% two extremes along the approximation continuum with \kwarg being entirely
% parsimony-based while \texttt{ARGinfer} is a CwR-based MCMC sampler.

%% Recent progress in ARGs through approximate structures}

% The second group of methods (\tsinfer, \relate, \argneedle) has marked a significant
% leap in scalability. These methods derive their scalability at least to a certain
% degree from the fact that the ARGs they infer are approximate structures. In particular,
% recombination is an emergent property of the inferred ARGs, explaining tree topology
% changes going from one local tree to the next. The recombination events themselves
% are not explicitly inferred and are thus not part of the resulting data structure.

% The computational gains from not having to pinpoint recombination events tie in
% with the inevitable uncertainty associated with ARG reconstruction [Hayman, 2022, ADD REFS].
% Some approaches have embraced the approximate nature of the resulting ARG by, for
% example, highlighting the inability to resolve certain splits and instead representing
% them as polytomies \citep{kelleher_inferring_2019}. Taken together this means that the
% degree of completeness in which genetic inheritance from ancestors to descendants is documented
% by each of these methods varies extensively.
% [TO ADD: these methods only provide point estimates]

\section{Data availability}

All scripts and data used for this manuscript are available on https://github.com/gertjanbisschop/smc-bit-paper
An implementation of the algorithm is available on https://github.com/gertjanbisschop/runsmc.
\FloatBarrier
\bibliographystyle{plainnat}
\bibliography{refs.bib,paper.bib}

\setcounter{secnumdepth}{2} % Print out appendix section numbers

\section*{Appendix}
\appendix

\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}

% % % % % % % % % % % % % % % % % %
\subsection*{Algorithm} \label{par:algo}
In this section we describe how to efficiently compute the likelihood under the SMC
from an ARG stored using the \texttt{tskit} tree sequence encoding. 
The tree sequence format encodes relationships using \emph{edges},
along with additional information that
allows us to sequentially recover (and reason about 
their differences) 
efficiently~\citep{kelleher_efficient_2016, ralph_efficiently_2020}.
The order of the edge deletions and insertions required to move from one local tree $T_1$ to
the next, $T_2$ imposes the strict total ordering on edges
(see definition \eqref{def:coal})
that is needed to efficiently maintain the state of $I_e(t)$ for each focal edge $e$.
Indeed, following the removal of edges that do not persist in $T_2$,
we simply declare that each subsequently added edge
can only coalesce with the already-visited edges 
that make up the (partially) reconstructed local tree $T_2$.

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{figures/ts_algo_2rows.png}
\caption{Likelihood computation algorithm. 
Moving along the genome we transition from
local tree $T_1$ to the next, $T_2$ in three steps. 
Edges that do not persist in $T_2$ are
removed first. Edges starting at the left coordinate of $T_2$ are inserted.
After deletion or insertion of an edge the count vector $I$ is updated along the
corresponding internode intervals. Following an insertion, and prior to updating $I$,
the likelihood for that edge is computed as $I$ then represents the total number of
lineages the lineage associated with this new edge can
coalesce with. By moving along the genome in this way, each edge gets visited exactly once.}
\label{fig:algo}
\end{figure}

This allows us to compute a likelihood in a single pass over all edges.
More concretely, the likelihood is computed by maintaining a single vector $I$ that tracks
the number of extant lineages in each internode interval (see Fig.~\ref{fig:algo}).
When moving from $T_1$ to $T_2$, the vector $I$ is first decremented along 
the internode intervals
spanned by the time of the child and the parent of all edges that do not persist in $T_2$.
For each subsequently inserted edge, we first compute the likelihood contribution
of that edge by computing the integral
in equation \eqref{eq:full-lik} using $I$, and then update the counts in $I$.
We further keep track of
the last parent a child node was connected to to detect recombination events.

% [TODO another paragraph summarising what you actually do]

% There is probably no need to potentially confuse people with this detail
%Remark: better to describe edges here as edge-intervals, each describing a
%single contiguous interval of genome inheritance between a pair of nodes.
%Under the SMC and when recording unary nodes all edges can be represented
%by a single edge-interval.

% ADD IN RESULTS:
% Figure \ref{sup:fig:vs-argweaver} shows the expected near-perfect correlation
% between the coalescent prior as returned by \argweaver and the likelihood as computed here.
% We simulated 1Mb of data for 10 diploid individuals
% under the SMC using human-like parameters. 100 samples were taking from the posterior every for
% a total of 1000 MCMC iterations.

% \subsection{Multiple segments description}
% 
% \comment{This is copied from above in case it's helpful in the future.}
% 
% We now define the process more formally,
% and following the notation of \citet{mcvean_approximating_2005}.
% At any point in time, the state of the process is $L(t)$,
% the set of labeled lineages extant at time $t$,
% where each ``lineage'' is represented by a union of non-overlapping half-open intervals
% which represent the ancestral material carried by that lineage,
% and labels are integers.
% So, we can write $L(t) = \{X_i(t)\}_{i=1}^{n(t)}$,
% where the $i^\text{th}$ lineage is $X_i(t)$
% and is of the form
% $X_i(t) = ([x_{i0}, y_{i0}), \dotsc, [x_{im_i}, y_{im_i}))$
% for some interval endpoints $x_{i0} < y_{i0} < x_{i1} < \cdot < y_{im_i}$.
% The \emph{span} of this lineage is the interval $s(X_i(t)) = [x_{i0}, y_{im_i}$,
% and its \emph{size} is $|X_i(t)| = y_{im_i} - x_{i0}$.
% Since we look backwards in time, $t=0$ is today, and so
% $L(0)$ consists of $n$ sampled lineages, labelled $0$ to $n-1$,
% each represented by a single interval spanning the entire genome.
% Then, the process evolves by a succession of coalescence and recombination
% events until each segment of ancestral material is only present in one lineage.
% The waiting time to next event is determined by these two competing processes
% with exponentially distributed waiting times as outlined below.
% 
% Recombination is described by a Poisson process of rate $r$ per unit of genomic length and time:
% so, if $T(t) = \sum_i |X_i(t)|$ is the sum of the total sizes of all lineages,
% then recombination occurs at instaneous rate $rT(t)$.
% A recombination to the left of $x$
% (i.e., between $x$ and $x-1$)
% with $x>x_{i0}$ and $x<y_{im}$ splits lineage $i$ into two new lineages,
% which each are given new, unique labels
% (and lineage $i$ is removed).
% This operation keeps the total amount of ancestral material unchanged.
% 
% Coalescence occurs between
% two lineages with overlapping spans at rate $\lambda = 1/(2N_e)$
% (if we assume a randomly mating, diploid population of constant size $N_e$).
% The newly formed lineage acquires the union of both ordered intervals, % L is updated accordingly
% and the original lineages are removed.
% Although coalescence is reciprocal, here we'll define a strict total order
% on all lineages in $L(t)$ based on their leftmost starting points (and index to break ties).
% For any such strict total order,
% the instantaneous rate of coalescence then equals $\lambda$ multiplied by the number of overlapping pairs,
% i.e., $\lambda \sum_{i \neq j} I_{ij}$,
% where
% \begin{equation}
% I_{ij} = \begin{cases}
%     1 & X_i > X_j \text{ and } s(X_i) \cap s(X_j) \neq \emptyset \\
%     0 & \text{otherwise.}
% \end{cases}
% \end{equation}
% We say that lineage $j$ can coalesce \emph{into} lineage $i$ if $I_{ij} = 1$
% (and notice that if $I_{ij} = 1$ then $I_{ji} = 0$).
% 
% For future use, define $C_i(t)$ to be the set of lineages that can coalesce into lineage $i$,
% i.e., $C_i(t) = \{X_j \in L(t) | I_{ij}(t) = 1\}$.
% (Therefore, $|C_i(t)| = I_{i}(t) = \sum_{j} I_{ij}(t)$.)
% Because $I_{ij}$ is defined in terms of a total order, the sets $C_i(t)$ are disjoint,
% which will simplify the likelihood computations.
% In particular,
% although recombination affects \emph{which} lineages can coalesce into $i$,
% it does not change the \emph{number} of such lineages:
% in other words, a recombination event changes $C_i(t)$ but not $I_{i}$.
% Another useful quantity to know is the number of lineages carrying material
% at each point on the genome:
% for a location $z$ and time $t$, this is
% \begin{equation}
%     N(t,z) = \#\{i : x \le z < y \text{ for some } [x,y) \in X_i(t) \} .
% \end{equation}
% Note that here we are not counting the ``gaps'' in a lineage,
% so that $N(t, z)$ might be less than the number of $i$ with $z \in s(X_i(t))$.
% 
% The recombination and coalescence operations described above
% either split a lineage or merge two overlapping lineages,
% and so as described so far, each lineage will consist of only a single interval,
% rather than a collection of intervals.
% However, there is one more operation:
% when a sub-segment of a lineage is ancestral to all samples,
% that sub-segment is removed.
% Formally, define $Z(t)$ to be the segments on which the lineages have totally coalesced:
% i.e., $Z(t) = \{z : N(t,z) = 1$.
% Then, each coalescence event replaces the resulting lineage $X_i(t)$
% by $X_i(t) \setminus Z(t)$,
% the set of intervals that results from removing $Z(t)$ from all the intervals in $X_i(t)$.
% 
% % note on the disjoint set of intervals for each lineage under the SMC
% Because of this last step, a single lineage can consist of more than one disjoint interval.
% As noted by \cite{mcvean_approximating_2005},
% this implies that the backwards-in-time description of the
% SMC is slightly different from the left-to-right description,
% in which single ancestors only ever span a single segment of genome.
% This is depicted in fig.\ref{fig:smc-unary}:
% for instance, node 8 is ancestral to all samples on the middle two trees.
% However, this does not affect the distribution
% of marginal genealogies up to the roots.
% More precisely, all lineages will be associated with
% a single interval up to the point where one section of the
% genome has reached its most recent common ancestor but the neighbouring parts have not.

\end{document}